---
title: "Mapping Data in R"
author: "Mehadi Hasan"
date: "2025-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Install sf package
```{r}
# CRAN
#install.packages("sf")

# MacOS
install.packages("sf", type = "binary")

# Install development versions from GitHub with:
#library(remotes)
#install_github("r-spatial/sf")
```

# Load Packages
```{r}
library(tidyverse)     # for data cleaning and plotting
library(gardenR)       # for Lisa's garden data
library(lubridate)     # for date manipulation
library(openintro)     # for the abbr2state() function
library(palmerpenguins)# for Palmer penguin data
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(gplots)        # for col2hex() function
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(ggthemes)      # for more themes (including theme_map())
theme_set(theme_minimal())
```

```{r}
# Starbucks locations
Starbucks <- read_csv("https://www.macalester.edu/~ajohns24/Data/Starbucks.csv")

# Lisa's favorite St. Paul places - used in leaflet example
favorite_stp_by_lisa <- tibble(
  place = c("Home", "Macalester College", "Adams Spanish Immersion", 
            "Spirit Gymnastics", "Bama & Bapa", "Now Bikes",
            "Dance Spectrum", "Pizza Luce", "Brunson's"),
  long = c(-93.1405743, -93.1712321, -93.1451796, 
           -93.1650563, -93.1542883, -93.1696608, 
           -93.1393172, -93.1524256, -93.0753863),
  lat = c(44.950576, 44.9378965, 44.9237914,
          44.9654609, 44.9295072, 44.9436813, 
          44.9399922, 44.9468848, 44.9700727)
  )

# North Carolina births and sids deaths 

nc <- st_read(system.file("shape/nc.shp", package="sf"), 
              quiet = TRUE) %>% 
  mutate(sid_per_1000birth_79 = SID79/BIR79*1000)

# Compute centroid (center) of each county
county_centroid_lat_long <- as.tibble(st_coordinates(st_centroid(nc))) %>% 
  rename(centr_long = X, 
         centr_lat = Y)

# County names and sid_per_1000birth_79
nc_centroid_county <- st_drop_geometry(nc) %>% 
  select(NAME, sid_per_1000birth_79)

# All centroid level info
nc_centroid <-  county_centroid_lat_long %>% 
  bind_cols(nc_centroid_county)
```

# Plotting points on a map
The Starbucks data, compiled by Danny Kaplan and provided by Alicia Johnson, contains information about every Starbucks in the world at the time the data were collected. It includes the Latitude and Longitude of each location. Let’s start by using familiar plotting tools

```{r}
ggplot(data=Starbucks) +
  geom_point(aes(x = Longitude, y = Latitude), 
             alpha = 0.2, 
             size = .1)
```

# Get the map information
```{r}
world <- get_stamenmap(
    bbox = c(left = -180, bottom = -57, right = 179, top = 82.1), 
    maptype = "terrain",
    zoom = 2)

# Plot the points on the map
ggmap(world) + # creates the map "background"
  geom_point(data = Starbucks, 
             aes(x = Longitude, y = Latitude), 
             alpha = .3, 
             size = .1) +
  theme_map()
```

